Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-24 introduction.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

There have been a number of numerical methods described in the
astronomical literature that contain elements of SAMR or have a
similar aim. For example, the N-body solver developed by
\citet{Villumsen89} used non-adaptive meshing to increase the
resolution in pre-selected regions. This static approach was later
used extensively when applied to hydrodynamics
\citep[e.g.,][]{Ruffert94, Anninos94}. Adding adaptivity is a more
recent enhancement, and there are now a number of N-body codes that
possess this feature, both with and without hydrodynamics
\citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97,
Truelove98, flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04,
Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}. Of these,
perhaps the most comparable and widely used are FLASH
\citep{flash_method}, which uses grid blocks of fixed size, and RAMSES
\citep{RAMSES} and ART \citep{ART97}, both of which refine individual
cells. It is also possible to deform the grid to obtain high
resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a
few grid codes have adopted an unstructured approach based on a moving
Voronoi mesh \citep{Arepo10, Tess11}.

In this paper, we present \enzo; a structured adaptive mesh refinement
(SAMR) code. Originally developed for cosmological hydrodynamics,
\enzo\ has since been used on a wide variety of problems.  It has
grown to become a general tool for astrophysical fluid dynamics and is
intended to be efficient, accurate and easily extended to include new
capabilities.  Although many of the components of the \enzo\ code have
been described in previous publications \citep{1995CoPhC..89..149B,
BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01,
Oshea04, 2007arXiv0705.1556N,
WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray},
there has previously been no systematic and complete description of
the code.  In this paper we provide that description, filling in many
previous omissions and showing the code's performance for a wide
variety of test problems.

The \enzo\ code has been extensively used over the last two decades in
a wide variety of problems, resulting in the publication of more than
100 peer-reviewed papers based on results obtained with the code. The
variety of astrophysical systems that \enzo\ has been used for include
galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L,
  2012ApJ...749..140H, 2013MNRAS.432.1989S}, galaxy
clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
\citep{Slyz05}, the intergalactic medium \citep{Fang01, 
2011ApJ...731....6S}, the circumgalactic
medium~\citep{2013MNRAS.430.1548H}, cooling flows \citep{Li12, 2013ApJ...763...38S},
turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the
formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
\citep{Collins11, 2011ApJ...727L..20K, Collins12a}.

Numerical simulations of astronomical phenomena now play a key role,
along with observations and analytic theory, in pushing forward our
understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01,
DecadalSurvey10}.  But along with this role comes responsibility.  We
believe that those developing simulation tools must fulfill two key
obligations: the first is to make those tools available to the
community as a whole, much in the way that astronomical data are now
regularly made publicly available.  The second is to document, test
and refine those methods so that they can be critically evaluated and
expanded upon by others.  Our public release of the \enzo\ code represents our
attempt to meet the first of these obligations; this paper represents
our attempt to meet the second.

>>>>

There have been a number of numerical methods described in the
astronomical literature that contain elements of SAMR or have a
similar aim. For example, the N-body solver developed by
\citet{Villumsen89} used non-adaptive meshing to increase the
resolution in pre-selected regions. This static approach was later
used extensively when applied to hydrodynamics
\citep[e.g.,][]{Ruffert94, Anninos94}. Adding adaptivity is a more
recent enhancement, and there are now a number of N-body codes that
possess this feature, both with and without hydrodynamics
\citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97,
Truelove98, flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04,
Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}. Of these,
perhaps the most comparable and widely used are FLASH
\citep{flash_method}, which uses grid blocks of fixed size, and RAMSES
\citep{RAMSES} and ART \citep{ART97}, both of which refine individual
cells. It is also possible to deform the grid to obtain high
resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a
few codes have adopted an unstructured approach based on a moving
Voronoi mesh \citep{Arepo10, Tess11}.

In this paper, we present \enzo, a structured adaptive mesh refinement
(SAMR) code. Originally developed for cosmological hydrodynamics,
\enzo\ has since been used on a wide variety of problems.  It has
grown to become a general tool for astrophysical fluid dynamics and is
intended to be efficient, accurate and easily extended to include new
capabilities.  Although many of the components of the \enzo\ code have
been described in previous publications \citep{1995CoPhC..89..149B,
BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01,
Oshea04, 2007arXiv0705.1556N,
WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray},
there has previously been no systematic and complete description of
the code.  In this paper we provide that description, filling in many
previous omissions and showing the code's performance for a wide
variety of test problems.

The \enzo\ code has been extensively used over the last two decades in
a wide variety of problems, resulting in the publication of more than
100 peer-reviewed papers based on results obtained with the code. The
variety of astrophysical systems that \enzo\ has been used for include
galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L,
  2013MNRAS.432.1989S}, galaxy
clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
\citep{Slyz05}, the intergalactic medium \citep{Fang01, 
2011ApJ...731....6S}, the circumgalactic
medium~\citep{2013MNRAS.430.1548H}, cooling flows \citep{Li12, 2013ApJ...763...38S},
turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the
formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
\citep{Collins11, 2011ApJ...727L..20K, Collins12a}.

Numerical simulations of astronomical phenomena now play a key role,
along with observations and analytic theory, in pushing forward our
understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01,
DecadalSurvey10}.  But along with this role comes responsibility.  We
believe that those developing simulation tools must fulfill two key
obligations: the first is to make those tools available to the
community as a whole, much in the way that astronomical data are now
regularly made publicly available.  The second is to document, test
and refine those methods so that they can be critically evaluated and
expanded upon by others.  Our public release of the \enzo\ code (which
can be found at \url{http://enzo-project.org}) represents our
attempt to meet the first of these obligations; this paper represents
our attempt to meet the second.

### 2013-06-24 ms.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

\begin{abstract}
This paper describes the open-source code \enzo, which uses 
block-structured adaptive mesh refinement to provide high spatial and
temporal resolution for modeling astrophysical fluid flows.  The code
is Cartesian and can be run in 1, 2, and 3 dimensions, and supports a
large variety of physics, including hydrodynamics, ideal and non-ideal
magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
of fluids and particles), primordial gas chemistry, optically-thin
radiative cooling of primordial and metal-enriched plasmas (as well as
some optically-thick cooling models), radiation
transport, cosmological expansion, and models for star formation and
feedback in a cosmological context.  In addition to explaining the
algorithms implemented, we present solutions for a wide range of test
problems, demonstrate the code's parallel scaling properties, and
discuss the \enzo\ collaboration's code development methodology.
\end{abstract}

>>>>

\begin{abstract}
This paper describes the open-source code \enzo, which uses 
block-structured adaptive mesh refinement to provide high spatial and
temporal resolution for modeling astrophysical fluid flows.  The code
is Cartesian, can be run in 1, 2, and 3 dimensions, and supports a
large variety of physics, including hydrodynamics, ideal and non-ideal
magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
of fluids and particles), primordial gas chemistry, optically-thin
radiative cooling of primordial and metal-enriched plasmas (as well as
some optically-thick cooling models), radiation
transport, cosmological expansion, and models for star formation and
feedback in a cosmological context.  In addition to explaining the
algorithms implemented, we present solutions for a wide range of test
problems, demonstrate the code's parallel scaling properties, and
discuss the \enzo\ collaboration's code development methodology.
\end{abstract}

### 2013-06-24 overview.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

The comoving evolution equations are equivalent to the fixed
coordinate version in the non-cosmological limit ($a = 1$, $\dot{a} =
0$).  We include the cosmological terms with the understanding that
\enzo\ is not restricted to cosmological applications -- while
historically the code was written with cosmological situations in mind
(e.g. galaxy clusters), in more recent years it has been used to
simulate a much broader range of astrophysical environments.

@@

Any collisionless components (e.g. dark matter and stars) are modeled
by N-body particles, whose dynamics are governed by Newton's equations
in comoving coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \grad \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles contribute to the gravitational potential through
Poisson's equation (Equation \ref{eq:potential}).

@@

The code can include either a homogeneous radiation background, 
or evolve an inhomogeneous
radiation field either by directly solving the radiative transfer
equation along rays or by solving a set of moment equations derived
from the radiative transfer equation. In comoving coordinates
\citep[e.g.,][]{Gnedin97} the radiative transfer equation reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{a_{\rm em}}{a} \, \hat{n} \cdot \grad I_\nu -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is specific
intensity of the radiation, with dimensions of energy per time $t$ per
solid angle $\Omega$ per unit area per frequency $\nu$.  The second
term represents the propagation of radiation, where the factor $a_{\rm
  em}/a$ accounts for cosmic expansion since the time of emission.
The third term describes both the cosmological redshift and dilution
of radiation.  On the right hand side, the first term captures
absorption ($\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$), and the
second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ from any
sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, \enzo\ can also solve
the angle-averaged radiative transfer equation, through a
flux-limited diffusion approximation with couplings to both the gas
energy and chemical number densities. The equations solved are
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \div\left(E_r \vecv \right) &=& 
  \div\left(D\grad E_r\right) -
  \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \div\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
$e_c$ is the internal energy correction due to photo-heating and
chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define $E_r$ through first assuming a fixed frequency
spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t)
\chi(\nu)$, and then defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
The quantity $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ is  the
total opacity, $\eta$ is the field of radiation sources, $\Gamma$ is
the radiation induced photo-heating rate, and $\Lambda$ is the
chemical cooling rate.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

\begin{equation}
\fcond= -\kappa_{\rm cond} \, {\hat \textbf{b}} (\hat{\textbf{b}} \cdot \grad T),
\end{equation}
where $\hat{\textbf{b}}$ is the unit vector denoting the direction of
the magnetic field.

@@

In this section we briefly describe the numerical methods which are
used to solve the equations outlined in Section~\ref{sec.equations}.
We proceed through the numerical methods in the same order as will be
used in Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that
there is a one-to-one correspondence between each of the following
overviews and the complete description provided in the later sections.
The goal in this section is to introduce the reader to the basic
principles of the methods without drowning them in detail.

@@

The primary feature of the \enzo\ code is its Adaptive Mesh Refinement
(AMR) capability, which allows it to reach extremely large dynamical
ranges with limited computational resources, opening doors to applications otherwise
closed by finite memory and computational time. Unlike moving mesh
methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
AMR (also referred to as \emph{structured} AMR, or SAMR) utilizes an adaptive
hierarchy of grid patches at varying levels of resolution.  Each
rectangular grid patch (referred to as a ``grid'') covers some region
of space requiring higher resolution than its \emph{parent grid},
and can itself become the parent grid to an even more highly resolved
\emph{child grid}.

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum timestep allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1 \leq
\Delta t_0$, which is the minimum of all the allowed timesteps for
those grids.  If there are more levels, we repeat this procedure until
the bottom level of the hierarchy has been reached.  Once there, we
continue advancing the grids on the lowest level until they have
``caught up'' to the coarser parent grids (i.e. $\sum \Delta t_l =
\Delta t_{l-1}$).  This procedure repeats itself until all grids have
been advanced by a total time of $\Delta t_0$ (see
Figure~\ref{fig:wcycle} for a schematic).

@@

\paragraph{Godunov MHD with Constrained Transport (MHD)}
This MHD method is second-order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second-order hyperbolic
solver of \citet{Li08a} and the divergence free AMR scheme of
\citet{Balsara01}.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
root grid on each timestep.  The advantage of using this method is
that it is fast, accurate, and naturally allows both periodic and
isolated boundary conditions for the gravity, choices which are very
common in astrophysics and cosmology.  On subgrids, we interpolate the
boundary conditions from the parent grid (either the root grid or some
other subgrid). The Poisson equation is then solved on every timestep
using a multigrid technique on one subgrid at a time. Aside from
self-consistently calculating the gravitational potential arising from
the baryon fields and particles in the simulation, there are also a
number of options for specifying static gravitational fields, and in
fact including self-gravity is optional.

@@

Collisionless matter (e.g. dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced through a single timestep using a drift-kick-drift
algorithm \citep{Hockney88} to provide second-order accuracy even in
the presence of varying timesteps.  Since the particles follow the
collapse of structure, they are not adaptively refined.  Nor are there
duplicate sets of particles for each level; instead, each particle is
associated with the highest refined level available at its position in
the domain and particles are moved between grids as the hierarchy is
rebuilt. Thus a particle has the same timestep and feels the same
gravitational force as a grid at that refinement level.

Although the particles are fixed in mass once initialized (with the
exception of star particles, which can lose mass in the feedback
process), we can create them with any initial set of masses and
positions.  For example, in many cosmological simulations static
subgrids are included from the beginning in order to improve the
initial spatial and baryonic mass resolution, and these subgrid are
populated with lower mass particles to correspondingly improve the
collisionless mass resolution.
%% One particle per initial grid point seems to provide approximately
%% equal sampling between the dark matter and gas. \mqk{This
%% contradicts the conclusions of O'Shea et al. 2005 that an initial
%% grid twice as fine as the mean particle separation is necessary to
%% properly capture the abundance of low mass halos.}

@@

\enzo\ can operate in a number of different modes with regard to
radiative cooling and heating. In the simplest mode, where the
multi-species flag is turned off and no individual chemical species
are tracked, the cooling rate is computed from a
simple temperature-dependent cooling function, taken from
\citet{SW87}.  If chemistry is turned on, then the code can include
cooling from all species of hydrogen and helium (including H$_2$) --
and the primordial cooling rates are computed in the same Jacobi iteration as the
chemistry.  It is also possible to include metal cooling based on a
set of multi-dimensional lookup tables computed with the
\textsc{Cloudy} code \citep{1998PASP..110..761F} as described in
\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}. Note that
the cooling and heating is most commonly treated in the optically-thin
limit, but the code can also follow radiative transfer in a limited
set of energy bins.

@@

A family of simple heuristic methods are used to model the formation
of stars and their feedback of metals and energy into the gas.  These
methods are based on the work of \citet{CO1992}, and involve the
identification of plausible sites of star formation based on a set of
criteria (for example, dense gas with a short cooling time, which is
both collapsing and unstable).  The local star formation rate is
computed using a range of methods, such as a density-dependent method
based on the Schmidt-Kennicutt relation \citep{K89}.  The affected gas
is converted into a star particle over a few dynamical times, and
metals and thermal energy are injected into the region surrounding the
star particle.  A related set of methods involves the simulation of
single Population III stars, rather than ensembles, and is calibrated
by \textit{ab initio} simulations of primordial star formation.

>>>>

The comoving evolution equations are equivalent to the fixed
coordinate version in the non-cosmological limit ($a = 1$, $\dot{a} =
0$).  We include the cosmological terms with the understanding that
\enzo\ is not restricted to cosmological applications -- while
historically the code was written with cosmological situations in mind
(e.g., galaxy clusters), in more recent years it has been used to
simulate a much broader range of astrophysical environments.

@@

Any collisionless components (such as dark matter and stars) are modeled
by N-body particles, whose dynamics are governed by Newton's equations
in comoving coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \grad \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles contribute to the gravitational potential through
Poisson's equation (Equation \ref{eq:potential}).

@@

The code can include either a homogeneous radiation background 
or evolve an inhomogeneous
radiation field either by directly solving the radiative transfer
equation along rays or by solving a set of moment equations derived
from the radiative transfer equation. In comoving coordinates
\citep[e.g.,][]{Gnedin97} the radiative transfer equation reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{a_{\rm em}}{a} \, \hat{n} \cdot \grad I_\nu -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is specific
intensity of the radiation, with dimensions of energy per time $t$ per
solid angle $\Omega$ per unit area per frequency $\nu$.  The second
term represents the propagation of radiation, where the factor $a_{\rm
  em}/a$ accounts for cosmic expansion since the time of emission.
The third term describes both the cosmological redshift and dilution
of radiation.  On the right hand side, the first term captures
absorption ($\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$), and the
second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ from any
sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, \enzo\ can also solve
the angle-averaged radiative transfer equation through a
flux-limited diffusion approximation, with couplings to both the gas
energy and chemical number densities. The equations solved are
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \div\left(E_r \vecv \right) &=& 
  \div\left(D\grad E_r\right) -
  \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \div\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
$e_c$ is the internal energy correction due to photo-heating and
chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define $E_r$ through first assuming a fixed frequency
spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t)
\chi(\nu)$, and then defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
The quantity $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ is  the
total opacity, $\eta$ is the field of radiation sources, $\Gamma$ is
the radiation-induced photo-heating rate, and $\Lambda$ is the
chemical cooling rate.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

\begin{equation}
\fcond= -\kappa_{\rm cond} \,\myvec{b} ( \myvec{b} \cdot \grad T),
\end{equation}
where $\myvec{b}$ is the unit vector denoting the direction of
the magnetic field.

@@

In this section we briefly describe the numerical methods that are
used to solve the equations outlined in Section~\ref{sec.equations}.
We proceed through the numerical methods in the same order as will be
used in Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that
there is a one-to-one correspondence between each of the following
overviews and the complete description provided in the later sections.
The goal in this section is to introduce the reader to the basic
principles of the methods without drowning them in detail.

@@

The primary feature of the \enzo\ code is its Adaptive Mesh Refinement
(AMR) capability, which allows it to reach extremely large spatial and
temporal dynamical
ranges with limited computational resources, opening doors to applications otherwise
closed by finite memory and computational time. Unlike moving mesh
methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
AMR (also referred to as \emph{structured} AMR, or SAMR) utilizes an adaptive
hierarchy of grid patches at varying levels of resolution.  Each
rectangular grid patch (referred to as a ``grid'') covers some region
of space requiring higher resolution than its \emph{parent grid},
and can itself become the parent grid to an even more highly resolved
\emph{child grid}.

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum timestep allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1 \leq
\Delta t_0$, which is the minimum of all the allowed timesteps for
those grids.  If there are more levels, we repeat this procedure until
the bottom level of the hierarchy has been reached.  Once there, we
continue advancing the grids on the lowest level until they have
``caught up'' to the coarser parent grids (i.e., $\sum \Delta t_l =
\Delta t_{l-1}$).  This procedure repeats itself until all grids have
been advanced by a total time of $\Delta t_0$ (see
Figure~\ref{fig:wcycle} for a schematic).

@@

\paragraph{Godunov MHD with Constrained Transport (MHD)}
This MHD method is second-order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second-order hyperbolic
solver of \citet{Li08a} and the divergence-free AMR scheme of
\citet{Balsara01}.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier technique \citep{Hockney88} to solve Poisson's equation on the
root grid on each timestep.  The advantage of using this method is
that it is fast, accurate, and naturally allows both periodic and
isolated boundary conditions for the gravity, choices which are very
common in astrophysics and cosmology.  On subgrids, we interpolate the
boundary conditions from the parent grid (either the root grid or some
other subgrid). The Poisson equation is then solved on every timestep
using a multigrid technique on one subgrid at a time. Aside from
self-consistently calculating the gravitational potential arising from
the baryon fields and particles in the simulation, there are also a
number of options for specifying static gravitational fields, and in
fact including self-gravity is optional.

@@

Collisionless matter (e.g. dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced through a single timestep using a drift-kick-drift
algorithm \citep{Hockney88} to provide second-order accuracy even in
the presence of varying timesteps.  Since the particles follow the
collapse of structure, they are not adaptively refined.  Nor are there
duplicate sets of particles for each level; instead, each particle is
associated with the highest refined level available at its position in
the domain and particles are moved between grids as the hierarchy is
rebuilt. Thus a particle has the same timestep and feels the same
gravitational force as a grid cell at that refinement level.

Although the particles are fixed in mass once initialized (with the
exception of star particles, which can lose mass in the feedback
process), we can create them with any initial set of masses and
positions.  For example, in many cosmological simulations static
subgrids are included from the beginning in order to improve the
initial spatial and baryonic mass resolution, and these subgrid are
populated with lower-mass particles to correspondingly improve the
collisionless mass resolution.
%% One particle per initial grid point seems to provide approximately
%% equal sampling between the dark matter and gas. \mqk{This
%% contradicts the conclusions of O'Shea et al. 2005 that an initial
%% grid twice as fine as the mean particle separation is necessary to
%% properly capture the abundance of low mass halos.}

@@

\enzo\ can operate in a number of different modes with regard to
radiative cooling and heating. In the simplest mode, where the
multi-species flag is turned off and no individual chemical species
are tracked, the cooling rate is computed from a
simple temperature-dependent cooling function, taken from
\citet{SW87}.  If chemistry is turned on, then the code can include
cooling from all species of hydrogen and helium (including H$_2$ and
Deuterium-related species such as HD) --
and the primordial cooling rates are computed in the same Jacobi iteration as the
chemistry.  It is also possible to include metal cooling based on a
set of multi-dimensional lookup tables computed with the
\textsc{Cloudy} code \citep{1998PASP..110..761F} as described in
\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}. Note that
the cooling and heating is most commonly treated in the optically-thin
limit, but the code can also follow radiative transfer in a limited
set of energy bins.

@@

A family of simple heuristic methods are used to model the formation
of stars and their feedback of metals and energy into the gas.  These
methods are based on the work of \citet{CO1992}, and involve the
identification of plausible sites of star formation based on a set of
criteria (for example, dense gas with a short cooling time, which is
both collapsing and unstable).  The local star formation rate is
computed using a range of methods, such as a density-dependent method
based on the Schmidt-Kennicutt relation \citep{K89}.  The affected gas
is converted into a star particle over a few dynamical times, and
metals and thermal energy are injected into the region surrounding the
star particle.  A related set of methods involves the simulation of
single Population III stars rather than ensembles, and is calibrated
by \textit{ab initio} simulations of primordial star formation.
